---
title: "ESM 206 Lab 10"
author: "An Bui"
date: "05 December 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Lab 10 Objectives:

- Correlation matrices
- Binary logistic regression
- Dates/times with lubridate
- Heatmap of solar irradiation in Santa Barbara, Hilo (HI), and Fairbanks (AK) in 2010
- Aquaculture with gganimate

# PART 0. Load libraries
```{r packages, include = FALSE}

library(tidyverse) # The tidyverse!
library(corrplot) # For correlation matrices
library(janitor) # For cleaning up column names
library(lubridate) # For dealing with dates & times
library(gganimate) # For adding animation to ggplot graphics
library(ggimage) # For updating graph components with images


```

# PART 1. Correlation matrix (World Bank world environmental factors data)

Compiled World Bank data, accessed from: https://www.kaggle.com/zanderventer/environmental-variables-for-world-countries#World_countries_env_vars.csv  

Create a correlation matrix to explore colinearity

- Read in the "world_env_vars.csv file" as env_var
- Find correlations between all variables, columns 2:28
- Plot the correlation matrix (corrplot)

```{r env_corr, echo = FALSE}
# read in csv and remove NA so that you can use cor()
# do complete case deletion: all rows with NA values are deleted
env_var <- read_csv("world_env_vars.csv") %>% 
  na.omit

# calculate Pearson's r for correlations
# find correlations between all columns in env_var except column 1 (country name)
cor_df <- cor(env_var[2:28])

# use corrplot() in corrplot package to make correlation matrix
corrplot(cor_df)
# size of circles and transparency indicate strength of correlation, color indicates directionality

# update corrplot to be easier to use
corrplot(cor_df,
         # only upper correlation matrix
         type = "upper",
         # ellipse direction indicates direction of correlation
         method = "ellipse",
         # make text black
         tl.col = "black",
         # space out text
         tl.cex = 0.5)

# more narrow = more strongly correlated
# right = negative correlation, left = positive correlation
# red = negative correlation, blue = positive correlation
# light = weak correlation, dark = strong correlation
```

# PART 2. Binary Logistic Regression (Donner Party Data)

Use the 'glm' function for fitting *generalized linear models* (the *logit* - log odds of survival, in our case, will be linearly related to Sex and Age. So we expect the final model to look something like this: 

$$Log Odds (Survival) = \beta_0 + \beta_1(Age) + \beta_2(Sex)$$

We'll use 'family = binomial' to run binomial logistic regression...otherwise, this looks very similar to other types of regression we've already done.  

**Question:** What effect do age and sex have on survivorship in the Donner party?

a. Read in the DonnerTable.csv file as DonnerTable

```{r donner_data, include = FALSE}
# read in csv
DonnerTable <- read_csv("DonnerTable.csv")
```

**Important:** When outcome dependent variable is already coded as 0/1, it's easy to define probability of outcome A or B. If you have a binary dependent variable is true/false, you need to figure out how R codes each as 0 or 1. The easiest way to know which is which is to specify on your own (using `r casewhen()`) so that there's no confusion about what you are defining as 0 or 1.  

b. Binomial logistic regression

```{r blr, echo = FALSE}
# use glm() to set up model
donner_blr <- glm(Outcome ~ Sex + Age, 
                  # choose the type of model you're constructing
                  family = "binomial",
                  # call data that you're using
                  data = DonnerTable)

summary(donner_blr)
# looking at intercepts, male survival is lower than female survival AND an increase in age leads to a decrease in the odds of survival
```

c. Use the model to answer: What are the log odds of survival for a 5 year-old female? The probability of survival?

```{r predict_f_5}

# 1. Create a data frame with variables Sex and Age, containing data "Female" and 5, respectively: 

# create a new data frame for a female of age 5
# data has to have the exact same variable names as the model you're using in order for things to match up
f_5 <- data.frame(Sex = "Female", Age = 5)


# 2. Find the log odds of survival for the new data (5 year old female) using predict() function with type = "link":

# give predict() model name
f_5_logodds <- predict(donner_blr, 
                       # new data frame
                       newdata = f_5,
                       # calculate log odds
                       type = "link")
f_5_logodds


# 3. Exponentiate the log odds to find ODDS of survival for a 5 year old female:

# figure out actual odds of survival for 5 year old female
f_5_odds <- exp(f_5_logodds)
f_5_odds
# odds are 4.23:1, odds of survival for 5 year old female ~80% BUT you can do this by writing in type = "response"

# Ask: Could we manually find the probability of survival for a 5-year old female? recall: p/(1-p) = ODDS



# 4. Actually, let's just use type = "response" in the predict function, which converts to a probability for us:

f_5_prob <- predict(donner_blr, 
                       # new data frame
                       newdata = f_5,
                       # calculate probability of survival
                       type = "response")
f_5_prob

```

d. What is the probability of survival for a 25 year-old male? 

```{r predict_m_25}

# Similarly:

m_25 <- data.frame(Sex = "Male", Age = 25) # Make a new data frame

m_25_prob <- predict(donner_blr, newdata = m_25, type = "response") # Find probability of survival
m_25_prob

```

e. Create new sequences of data so that we can graph probabilities for the entire spectrum of ages, designated by sex.

```{r new_data}

seq_age <- rep(seq(from = 0, to = 100), 2) # Create a sequence from 0 to 100, twice (one will be "Male" and one will be "Female")

f_101 <- rep("Female", 101) # Repeat 'Female' 101 times (to match years data)
m_101 <- rep("Male", 101) # Repeat 'Male' 101 times
mf_101 <- c(f_101, m_101) # Combine them into a single vector

# Combine the age and sex sequences into a single data frame - that will be the new data that we have our model make predictions for

donner_newdata <- data.frame(seq_age, mf_101) # MUST make column names match variables in the model!
colnames(donner_newdata) <- c("Age","Sex")

```

f. Now that we have new data to put into our model to have it make predictions, let's go ahead and actually find the predicted probabilities for each Age/Sex combination.

```{r donner_predict}

# Find probabilities using predict (with type = "response"). Include SE.


# Coerce outcome into data frame. 



```

g. Graph results.

```{r donner_graph}

  
```


# PART 3. Solar irradiation at the 3 locations in 2010 (pull in all together, do some lubridate stuff, etc.)

a. Read in multiple solar irradiation files (for SB (CA), Hilo (HI), and Fairbanks (AK)):
```{r solar_data}




```

b. Wrangle the data
```{r wrangle_solar}



```

c. Use lubridate() functions to convert to times/dates
```{r date_time}


```

d. Make an awesome figure of solar irradiation (heat/tile map)
```{r solar_graph}



```


#PART 4. gganimate example: total aquaculture production (metric tons) for United States, Brazil, Chile, and Ecuador

a. Get the data, and do some wrangling:
```{r aqua_data}


```


b. Read in the fish.png as 'fish'

```{r fish_image}


```

c. Make a graph...with gganimate!
```{r gganimate_fish}


```

##END LAB